---
title: 'put the title of your project here'
author: "put your name here"
date: "Due Oct. 30, 2017"
bibliography: "decarbonization.bib" # this line tells RMarkdown which file to use for the bibliography
nocite: |
  @pielke_climate_fix_2010,
  @pielke_evaluation_2011,
  @pielke_british_2009,
  @pielke_mamizu_2009
output: 
  html_document: default
  pdf_document:
    include:
      in_header: "ees3310.sty" # This line just tells RMarkdown to use a style file to format the PDF file with Times Roman font.
---

```{r setup, include=FALSE, cache = FALSE}
knitr::knit_hooks$set(inline = function(x) { knitr:::format_sci(x, 'md')})
knitr::opts_chunk$set(cache = TRUE, echo = TRUE, fig.height=4, fig.width=7)

# This section loads necessary R libraries and sources scripts that define 
# useful functions format_md.
# 
data_dir = "data"
script_dir = "scripts"

if (!dir.exists('data')) dir.create('data')

library(pacman)

# Some lab computers fail to load tidyverse
# because bindr is not installed.
# This will prevent such errors.
if (!p_isinstalled(bindr)) p_install(bindr)


# Load all of the following packages.
# Install any missing packages.
p_load(scales, tidyverse, stringr, janitor, knitr)
p_load_gh("gilligan-ees-3310/kayadata")

# This sets the default style for ggplot
theme_set(theme_bw(base_size = 15))

# Load basic utility scripts
source('scripts/utils.R', chdir = T)
```
# Top-Down Analysis of Decarbonization
```{r load_world_data}
target_year = 2050
reference_year = 2005
start_year = 2016

target_world_reduction = 0.36

world_data = get_kaya_data("World")
world_fuel_mix = get_fuel_mix("World") %>% arrange(fuel)
world_projection = project_top_down("World", target_year)

target_world_emissions = filter(world_data, year == reference_year)$F * (1 - target_world_reduction)
projected_world_E = world_projection$E
```
The pathway for keeping warming below 2&deg;C implies a target of reducing 
global CO~2~ emissions by `r format_md(100 * target_world_reduction, 0)`% 
below the `r reference_year` level by `r target_year`, which would be
`r format_md(target_world_emissions,1)`  million metric tons.
Top-down projections predict that global energy demand will be 
`r format_md(projected_world_E, 0)` quads in `r target_year`, so what does that imply for 
 the need to install low-carbon energy?
 
 The current energy mix is shown in the Figure below:
 
```{r world_fuel_mix_figure, fig.width=7, fig.height=5}
 e_factors = emissions_factors()
 
 plot_fuel_mix(world_fuel_mix)
 ```
 
Let's calculate what the emissions in `r target_year` would be if the world 
were still using this fuel mix.

```{r world_emissions_factor_table}
kable(e_factors)
```

First, let's calculate how many quads of energy of each fuel the world would
be consuming in 2050:

```{r world_fuel_mix_2050}
world_fuel_mix_2050 = world_fuel_mix %>% 
  select(fuel, pct) %>% # We only need to worry about the percentage of each fuel
  mutate(quads = projected_world_E * pct / 100) # We multiply the total energy demand by 
                                    # The percentage from each fuel to figure
                                    # the number of quads of each fuel that
                                    # the world would consume if the fuel
                                    # mix stayed the same.
  kable(world_fuel_mix_2050, digits = c(0,1,0))
```

Now, let's combine the emissions factors and the fuel mix into a single table:

````{r world_emissions_tab, warning = FALSE}
world_fuel_mix_2050 = world_fuel_mix_2050 %>% 
  left_join(e_factors, by = "fuel")     # left_join combines two tables, taking all 
                                        # rows from world_fuel, and combining the
                                        # rows of e_factors by matching the column 
                                        # "fuel" for the two tables. Because it's a 
                                        # left join, it ignores any rows of e_factors 
                                        # if the "fuel" doesn't have a matching row 
                                        # in world_fuel_mix
  
  kable(world_fuel_mix_2050, digits = c(0, 0, 0, 0, 1))  
```

Now that we combined the tables, we can multiply the number of quads by the 
emission factors to get the emissions from each fuel:

```{r world_emissions_2050}
# Multiply quads times emissions factor to get emissions
world_emissions_2050 = world_fuel_mix_2050 %>%
  mutate(emissions = quads * emission_factor)

# Print a table 
kable(world_emissions_2050, digits = c(0, 1, 0, 1, 0), 
      format.args = list("big.mark" = ","),
      col.names = c("Fuel", "%", "Quads", "MMT/Quad", "MMT CO2"))
```
```{r calculate_world_emissions_reduction}

# Add up all the emissions to get the total
total_world_emissions_2050 = sum(world_emissions_2050$emissions)

# Emissions reduction necessary to meet target
delta_world_emissions = total_world_emissions_2050 - target_world_emissions

# Separate each fossil fuel to make calculations easier

world_coal_2050 = filter(world_emissions_2050, fuel == "Coal")
world_gas_2050 = filter(world_emissions_2050, fuel == "Natural Gas")
world_oil_2050 = filter(world_emissions_2050, fuel == "Oil")

# Reduction from eliminating both coal and gas

world_coal_gas_reduction = world_coal_2050$emissions + world_gas_2050$emissions


# What's left that we have to meet by reducing oil consumption
world_remainder = delta_world_emissions - world_coal_gas_reduction

oil_factor = filter(e_factors, fuel == "Oil")$emission_factor

# Quads of oil we have to replace with renewables.
delta_world_oil_quads = world_remainder / oil_factor

# Total clean energy needs
world_new_clean_energy = delta_world_oil_quads + world_coal_2050$quads +
  world_gas_2050$quads
```

The total emissions in `r target_year` woulld be 
`r format_md(total_world_emissions_2050, digits = 0, comma = TRUE, scientific = FALSE)` 
million metric tons.
We need to get this down to the target of 
`r format_md(target_world_emissions, digits = 0, comma = TRUE, scientific = FALSE)`
million metric tons. Thus, we have to replace 
`r format_md(delta_world_emissions, digits = 0, comma = TRUE, scientific = FALSE)` 
million metric tons
of fossil-fuel emissions with renewable energy.

We can start by eliminating coal, which will reduce emissions by 
`r format_md(world_coal_2050$emissions, digits = 0, comma = TRUE, scientific = FALSE)`
MMT. 
Next, we can eliminate gas, which will reduce emissions by 
`r format_md(world_gas_2050$emissions, digits = 0, comma = TRUE, scientific = FALSE)`

(**Why do we eliminate gas next, when oil is responsible for much more emissions?**)

This gives us total emissions reduction of 
`r format_md(world_coal_gas_reduction, 0, comma = TRUE, scientific = FALSE)` MMT.
That leaves 
`r format_md(delta_world_emissions - world_coal_gas_reduction, 0, comma = TRUE, scientific = FALSE)`
MMT to be eliminated by reducing oil consumption.

Eliminating coal requires 
`r round(world_coal_2050$quads)` 
quads of clean energy to replace coal, 
eliminating gas requires
`r round(world_gas_2050$quads)` 
quads of clean energy to replace natural gas,
and finally, we must replace 
`r round(delta_world_oil_quads)` 
quads of oil with clean energy. 
This adds up to 
`r round(world_new_clean_energy)`
quads of new clean energy.


## Clean power requirements

If we want to provide this `r round(world_new_clean_energy)` quads per year
of clean 
energy using nuclear power, we need to figure out how many nuclear power
plants it takes to produce a quad per year.

Power plants have a "nameplate capacity", which indicates the power they
would produce if they ran full out 24/7. In fact, power plants can't do this
because they need down time for maintenance and solar and wind are also subject
to the fact that the sun does not shine all the time, nor does the wind blow
at its full strength all the time.

We represent the down-time by a "capacity factor," so the actual average power
a plant produces during a year is the product 
$$\text{average power} = \text{nameplate capacity} \times \text{capacity factor}$$
The following table shows the power output from different kinds of power plants.

```{r power plant capacity}
capacity = generation_capacity()
# Capacity has columns for:
# nameplate_capacity, which is the power the plant would produce if it ran

capacity = capacity %>%
  mutate(average_power = nameplate_capacity * capacity_factor)

kable(capacity)
```

Let's convert this to quads:

```{r power plant capacity_quads}
capacity = capacity %>% select(fuel, description, average_power) %>%
  mutate(quads = average_power / megawatts_per_quad(),
         plants_per_quad = 1 / quads)

kable(capacity, digits = c(0,0,2,5,1))
```

Now we are ready to calculate the number of power plants we have to construct between now and `r target_year`.

# References
